---
title: "HW7 - Exercise7"
author: JooChul Lee
date: "`r format(Sys.time(), '%d %B %Y')`"
documentclass: article
knit: "bookdown::render_book('Exercise2.Rmd', 'bookdown::pdf_document2')"
fontsize: 11pt
papersize: letter
header-includes:
   - \usepackage{algorithm}
   - \usepackage{algpseudocode}
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
## for latex and html output
isHtml <- knitr::is_html_output()
isLatex <- knitr::is_latex_output()
latex <- ifelse(isLatex, '\\LaTeX\\', 'LaTeX')

## specify global chunk options
knitr::opts_chunk$set(fig.width = 5, fig.height = 4, dpi = 300,
                      out.width = "90%", fig.align = "center")

```

# Normal mixture revisited

Under the assumption, $\delta$ is an uninformative prior and we use a simple random walk. The prior for $\mu_1$ and $\mu_2$ are $N(0, 10^2)$, that the prior for $\sigma_1^2$ and $\sigma_2^2$ are $IV\Gamma(a, b)$ with shape $a = .5$ and scale $b = 10$,$\pi(\delta), \pi(\mu_1), \pi(\mu_2) , \pi(\sigma^2_1)$, and $\pi(\sigma^2_2)$, respectively.

Suppose that we consider finite mixture normal distribution, $\bold{X}$

$$f(x) = \delta N(\mu_1,\sigma^2_1) + (1-\delta) \delta N(\mu_2,\sigma^2_2) $$

$f(x,\delta,\mu_1,\mu_2,\sigma^2_1,\sigma^2_2) \varpropto (\delta \frac{1}{\sigma_1}exp\{-\frac{(x -\mu_1)^2}{2\sigma^2_1}\} + (1-\delta) \frac{1}{\sigma_2}exp\{-\frac{(x -\mu_2)^2}{2\sigma^2_2}\})\times \pi(\delta) \times \pi(\mu_1) \times \pi(\mu_2) \times \pi(\sigma^2_1) \times \pi(\sigma^2_2)$

 
Then, MCMC using Gibbs sampling is the following;

Step 1 : Draw $\delta^t$ from $\pi(\delta^t| \mu^{t-1}_1, \mu^{t-1}_2 , \sigma^{t-1}_1,\sigma^{t-1}_2,x^t)$

Step 2 : Draw $\mu^t_1$ from $\pi(\mu_1| \delta^t, \mu^{t-1}_2 , \sigma^{t-1}_1,\sigma^{t-1}_2,x^t)$

Step 3 : Draw $\mu^t_2$ from $\pi(\mu_2| \delta^t, \mu^{t}_1 , \sigma^{t-1}_1,\sigma^{t-1}_2,x^t)$

Step 4 : Draw $\sigma^{t}_1$ from $\pi(\mu_2| \delta^t, \mu^{t}_1 , \mu^{t}_2,\sigma^{t-1}_2,x^t)$

Step 5 : Draw $\sigma^{t}_2$ from $\pi(\mu_2| \delta^t, \mu^{t}_1 , \mu^{t}_2,\sigma^{t}_1,x^t)$

Repeat Step 1 - 5

```{r}
library(MCMCpack)
library(HI)
delta <- 0.7
n <- 100
set.seed(254)
u <- rbinom(n, prob = delta, size = 1)
x <- rnorm(n, ifelse(u == 1, 7, 10), 0.5)

loglik = function(x,mu1,mu2,sigma1,sigma2,delta)
{
   Fisrt =  log( ( delta * dnorm(x, mu1,sigma1) ) + ( (1-delta) * dnorm(x,mu2,sigma2) ) )  
   return(sum(Fisrt))
}

HW7 = function(x,ini.D,ini.mu1,ini.mu2, ini.sigma1,ini.sigma2, iter)
{
   inix=x
   current.D <- ini.D; current.mu1 <- ini.mu1; current.mu2 <- ini.mu2; current.sigma1 <- ini.sigma1; 
   current.sigma2 <- ini.sigma2
   new.D <- new.mu1 <- new.mu2 <- new.sigma1 <- new.sigma2 <- rep(0,iter)
   for(i in 1:iter)
   {   
      D_v =function(delta) loglik(x =inix, mu1 = current.mu1, mu2 = current.mu2, sigma1 = current.sigma1, 
                                  sigma2 = current.sigma2, delta = delta)
      new.D[i] =  arms(current.D, D_v, function(delta) {(delta>0)*(delta<1)}, 1)

      mu1.v = function(mu1) loglik(x =inix, mu1 = mu1, mu2 = current.mu2, 
                             sigma1 = current.sigma1, sigma2 = current.sigma2, delta = new.D[i]) + 
         log( dnorm(mu1,0,10^2) )
      new.mu1[i] = arms(current.mu1, mu1.v, function(mu1) (mu1 > -30) * (mu1 < 30), 1)
   
      mu2.v = function(mu2) loglik(x =inix, mu1 = new.mu1[i], mu2 = mu2, 
                             sigma1 = current.sigma1, sigma2 = current.sigma2, delta = new.D[i]) + 
         log( dnorm(mu2,0,10^2) )
      new.mu2[i] = arms(current.mu2, mu2.v, function(mu2) (mu2 > -30) * (mu2 < 30), 1)
   
      sigma1.v = function(sigma1) loglik(x =inix, mu1 = new.mu1[i], mu2 = new.mu2[i], 
                             sigma1 = sigma1, sigma2 = current.sigma2, delta = new.D[i]) + 
         log(dinvgamma(sigma1, 0.5 , 0.1))
      new.sigma1[i] = arms(current.sigma1, sigma1.v, function(sigma1) (sigma1> 0) * (sigma1 < 100), 1)

      sigma2.v = function(sigma2) loglik(x =inix, mu1 = new.mu1[i], mu2 = new.mu2[i], 
                                   sigma1 = new.sigma1[i], sigma2 = sigma2, delta = new.D[i]) + 
         log(dinvgamma(sigma2, 0.5 , 0.1))
      new.sigma2[i] = arms(current.sigma2, sigma2.v, function(sigma2) (sigma2 > 0) * (sigma2 < 100), 1)

      current.D = new.D[i]; current.mu1 = new.mu1[i]; current.mu2 = new.mu2[i]; 
      current.sigma1 = new.sigma1[i]; current.sigma2 = new.sigma2[i]
   }
  list(new.D = new.D, new.mu1= new.mu1, new.mu2 = new.mu2, new.sigma1= new.sigma1, new.sigma2 = new.sigma2)
}
joo = HW7(x,0.5,1,1,1,1, iter = 3000)
plot(ts(joo$new.D[-(1:500)]), ylab = 'Delta')
hist(ts(joo$new.D[-(1:500)]), main = 'Delta',xlab = 'Delta')
plot(ts(joo$new.mu1[-(1:500)]), ylab = 'mu1')
hist(ts(joo$new.mu1[-(1:500)]), main = 'mu1',xlab = 'mu1')
plot(ts(joo$new.mu2[-(1:500)]), ylab = 'mu2')
hist(ts(joo$new.mu2[-(1:500)]), main = 'mu2',xlab = 'mu2')
plot(ts(joo$new.sigma1[-(1:500)]), ylab = 'sigma1')
hist(ts(joo$new.sigma1[-(1:500)]), main = 'sigma1',xlab = 'sigma1')
plot(ts(joo$new.sigma2[-(1:500)]), ylab = 'sigma2')
hist(ts(joo$new.sigma2[-(1:500)]), main = 'sigma2',xlab = 'sigma2')
```
500 runs are droped for the burn-in period.